Improving the Image Acquisition Rate \\ of an Atomic Force Microscope \\ Through Sub-sampling and Reconstruction

### 2019-06-25 latex/afm-cs-main.tex
Update on Overleaf.

\author{Roger A. Braker, Yufan Luo, Lucy Y. Pao and Sean B. Andersson
  \thanks{The authors are with the Dept. of Electrical, Computer, and
    Energy Engineering at the University of Colorado, 425 UCB,
    Boulder, CO 80309, United States. Phone: +1 (303) 492-2360. Fax:
    +1 (303) 492-2758. R. A. Braker (corresponding author
    roger.braker@colorado.edu) is a graduate student and L.Y. Pao
    (pao@colorado.edu) is the Palmer  Professor.}
  \thanks{This work was supported in part by the US National Science
    Foundation (NSF Grant CMMI-1234980)  Agilent Technologies,
    Inc., and the Hanse-Wissenschaftskolleg Institute for Advanced Study in Delmenhorst, Germany.} }

\maketitle
\begin{abstract}
  Undersampling-based approaches in Atomic Force Microscopy (AFM) aim
  to reduce the time to acquire an image by reducing the number of
  measurements needed while still maintaining image quality. In this
  paper, we describe a hardware implementation and demonstration of
  one such approach based on the use of collections of short scans
  known as $\mu$--paths. Using a commercial AFM, we acquire data on a
  grating sample. Reconstructions are made from these data using
  % three approaches: (1) inpainting, an interpolation technique that
  % diffuses information from sampled locations into unsampled
  % regions, (2)
  % basis-pursuit, a compressive sensing-based algorithm that seeks to
  % minimize a
  % measure of sparsity in the underlying image, and (3)
  a new variant of basis pursuit, Basis Pursuit with Vertical
  Variation (BPVV), that is designed to reduce artifacts arising from
  the sampling pattern. The quality of the resulting images is
  compared to images from standard raster scans of the same regions at
  comparable imaging rates using both the peak signal-to-noise ratio
  and the structured similarity index metric and a new metric we call
  the relative damage index. \todo{Re-write abstract, especially
    conclusions} These experiments demonstrate that at slow scan
  rates, the $\mu-$path scheme produces similar image quality with
  significantly less sampling (and thus less tip-sample interaction)
  and that at high rates, the undersampling scheme produces higher
  quality images than raster scanning.
\end{abstract}

@@

In addition to imaging static samples, AFM is used to study dynamics
in systems with nanometer-scale features
\cite{Yang:2017im,Shibata:2017da,Shibata:2015jd,Ando:2014ja}. The
image acquisition time of conventional AFMs, however, is typically on
the order of seconds to minutes, severely limiting the time-scales
that can be explored. Driven by the need for faster imaging, there
continue to be many active research efforts to overcome this
challenge. Since an AFM is fundamentally a mechanical microscope, many
of the approaches to high-speed AFM (HS-AFM) have focused on
modifications to the physical components. These have included the use
of small, fast cantilevers \cite{Viani:1999tp,
  braunsmann2010high,Adams:2016hg}, the development of faster
actuators \cite{Maroufi:2015gt,Yong:2012kd,Kenton:2012cm}, as well as
the application of advanced controllers
\cite{Rana:2018es,Yong:2015gr,butterworth2010adaptive,salapaka2002high}.
By combining different techniques, there are current generation,
high-end instruments that can image at rates on the order of 1-10
frames/sec \cite{Ando:2014ja}. However, the fastest rates are achieved
only over small scan ranges and there remain many systems of interest
whose dynamics are faster even than these instruments can achieve. In
addition, there is a large installed base of much slower AFMs that can
benefit from alternative approaches to improving imaging rate.

A complimentary class of HS-AFM approaches considers modifications to
the sampling scheme rather than the system mechanics. Alternative scan
trajectories like spirals, cycloids, and Lissajous figures, have been
used in place of the standard raster scan. These trajectories are
easier for the actuators to follow, allowing the tip to be moved more
quickly and leading to faster image acquisition on an otherwise
unmodified instrument
\cite{Bazaei:2017dm,Wu:2015dt,Rana:2014bj,Tuma:2012hv,Yong:2010gm}.
Such approaches are ultimately limited by the vertical bandwidth of
the instrument for regulating the deflection signal \cite{Teo:2016ev}.
Related to these are local scanning methods that use the measurements
in real-time to steer the tip to focus the scan on features of
interest, reducing imaging time by reducing the amount of sampling
needed \cite{Hartman:2018ja,Wen:2018fl,Zhang:2015cf,Huang:2014dw}.
While these have been shown to yield an order of magnitude or better
improvement in imaging rate, they are limited in the class of samples
that can be imaged.

@@

One simple way to create undersampling schemes is by modifying
existing full scan patterns including raster, spiral, and Lissajous
scanning. For example, subline sampling is generated by randomly
skipping some of the horizontal lines in raster scanning
\cite{han2018reconstruction,chen2013enhancement}. For spiral and
Lissajous patterns, the scan parameters (frequency and amplitude) can
be selected to ensure the trajectory only passes through a desired
fraction of the pixels in the final image. The scanning time for these
smooth undersampling patterns can be estimated based on the proportion
of the pixels in the trajectory. However, results in the literature of
CS make clear that \textit{randomness} in the sampling pattern is
essential for creating good reconstructions in the general case. The
smooth nature of these patterns reduces their randomness and thus
leads to less accurate reconstructions than a completely random
sampling of the pixels for the same sampling fraction.

In AFM, implementing a truly random pattern requires that the tip be
engaged with the surface to collect a measurement, lifted and moved to
the new location, and engaged again.
Because the re-engagement process
is typically slow, this can lead to excessively long image acquisition
times and negate the gains from undersampling \cite{andersson2012non}.
In \cite{maxwell2014compressed}, we introduced a discrete undersampling scheme,
called a $\mu$-path pattern, which consists of short randomly placed,
horizontal scans (see Fig.~\ref{fig:mu_mask} for one such pattern). It
is designed to balance the randomness needed (to ensure good
reconstruction) with continuous scanning (to reduce the number of tip
engagements).

Our previous work on the $\mu$-path pattern, using theoretical
calculations, simulations, and a preliminary implementation,
demonstrates that the approach can achieve scanning time
reduction while maintaining faithful image reconstruction
\cite{maxwell2014compressed,Luo:2015tu, braker_hardware_2018}. 
The main contributions of this paper relative to that past work are:
\begin{itemize}
\item We introduce a new reconstruction algorithm which is designed to
minimize the artifacts arising from the structure of the $\mu$-path
scanning pattern. This algorithm is described and compared to existing
reconstruction methods through simulations in Sec. \ref{sec:reconstructionMethods}.
\item In our prior experimental prototype \cite{braker_hardware_2018}, we used very basic integral controllers for all three axes for both raster and $\mu$-path scans. Here, we design compensators for each axis using standard loop-shaping techniques (Sec.~\ref{sec:control}), which result in a substantial increase in closed-loop bandwidth (e.g., 50 Hz to 450~Hz for the $Z$-axis).
We show in Sec.~\ref{sec:ff_control} that one of the limitations of $\mu$-path scanning is the discontinuous $X$-axis reference (e.g., in moving between $\mu$-paths) which excites the cross-coupling modes between the $X$ and $Z$ axes, an affect which neither our prior experimental work nor the numerous simulation studies have accounted for. We mitigate the cross coupling with a feed-forward design.
\item We experimentally compare (Sec. \ref{sec:results}) $\mu$-path scanning to not only full density raster scans but also to coarse raster scans which are interpolated to a full image. This is in contrast to typical sub-line sampling which takes scan lines at random. The clear advantage to a coarse raster scan is that it requires no modification to the instrument, only to the post-processing.
\item The bulk of this work is based on experiment, not simulation, in contrast to prior work by ourselves \cite{Luo_nano_2015,maxwell_acc_2014} and others \cite{han_optimal_2018,jensen_reconstruction_2013,oxvig_structure_2017}. Our main focus here is on delineating the practical issues which must be addressed to effectively realize $\mu$-path scanning, which has only seen limited attention in the past.
\end{itemize}

@@

\noindent
\begingroup \setlength{\tabcolsep}{1pt}
\begin{figure*}[ht!]
  \centering
  \begin{tabular}{cccccc}
    \textit{\small original image} & \textit{\small BPDN} & \textit{\small BPVV}
    & \textit{\small detail of original} & \textit{\small detail of BPDN} & \textit{\small detail of BPVV} \\
    \includesvg[width=0.16\textwidth]{figures/cs20ng_og.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_og_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp_zoom.svg}
    %
    &   \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv_zoom.svg}\\
    % 
    % ------------------------------
    % Row
    % 2
    % --------------------------------------------
    \includesvg[width=0.16\textwidth]{figures/dna_og.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bp.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_og_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bp_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bptv_zoom.svg}
  \end{tabular}
  \caption{Reconstruction comparison between BPDN and BPVV. (first
    column) Original raster-scanned image. (second column) BPDN
    reconstruction from random, 40 pixel long $\mu$-paths with 25\%
    sampling. (third column) BPVV reconstruction from the same
    sub-sampled data. (remaining columns) Corresponding details from
    the red boxes indicated in the raster image. The results show that
    BPVV reduces the artifacts arising from the horizontal scans of
    the $\mu$-path pattern that appear in BPDN reconstructions.}
  \label{fig:BPTV_demonstration}
\end{figure*}
\endgroup

In the AFM application, the probe can only measure a single pixel at a
time. Thus the rows of $R$ are a subset of the rows of an $n$ by $n$
identity matrix. Ideally, the sparsity basis and the measurement
matrix $R$ will have a low \textit{mutual coherence}, a measure of how
each of the rows of $R$ (the measurements) ``spreads out'' in the
domain of $U$ \cite{candes2007sparsity}. In the following, we assume
that $U^TU=I$, though in the simulation and experimental results, we
take $U^T$ as the Discrete Cosine Transform (DCT). The DCT basis
generally keeps a good balance between achieving a low mutual
coherence between $U$ and the $\mu$-path sensing matrices $R$ and
providing a high sparsity for typical AFM sample images.

Basis pursuit with denoising (BPDN) is one common realization of the
CS-based reconstruction problem, and is given by the optimization
\begin{equation}
  \min_{x} \left \| U^Tx \right \|_1 \quad
  \text{s.t.}\quad x\in Q_p = \{x:~\|Rx - y\|_2 < \sigma\} \label{op:bp}
\end{equation}
where $\sigma$ represents uncertainty in the measurements. BPDN
essentially searches for the sparsest signal from all the candidates
that match the measurements.

@@

To mitigate these affects, we propose a new variant of BPDN that we
term basis pursuit with vertical variation (BPVV). BPVV adds a
vertical total variation penalty in the spatial domain to the
optimization objective. That is, \eqref{op:bp} is modified to \todo{It
  has been my observation that some benefit is gained by including
  $D_h$, but weighting it less than $D_v$. Not sure if we should leave
  it in. If so, BPVV should be BPTV or something.}
\begin{align}
  \min_{x\in Q_p}~f(x) &= \min_{x\in Q_p} \left\|U^Tx\right\|_1
                         + \alpha\left\|D_vx\right\|_1
                         + \beta\left\|D_hx\right\|_1 \label{op:bp_tv}\\
                       &=\min_{x\in Q_p} \left\|W^Tx\right\|_1 \label{op:bp_tvW}
                       % \quad W
                       % = \begin{bmatrix}U&\alpha_hD^T_h&\alpha_vD^T_v \end{bmatrix}^T
                       % \nonumber
\end{align}
where
\begin{equation*}
  W = \begin{bmatrix}U^T&\alpha D_h&\beta D_v \end{bmatrix}
\end{equation*}
with ${D_v x=\text{vec}(\nabla_v X)}$ (resp.,
${D_v x=\text{vec}(\nabla_v X)}$) representing a discrete gradient of
a matrix in the vertical (resp., horizontal) direction (see, e.g.,
\cite[Sec. 6.1,]{becker_nesta_2011}).
% and
% \begin{align*}
%   (\nabla_v X)_{i,j} = 
%   \begin{cases}
%     X_{i+1,j} - X_{i,j}, & 1\leq i<h,~ 1\leq j\leq s,\\
%     0, & \phantom{:}i=h,~\phantom{:::::} 1\leq j\leq s.
%   \end{cases}\\
%   (\nabla_h X)_{i,j} =
%   \begin{cases}
%     X_{i,j+1} - X_{i,j}, & 1\leq i\leq h,~1\leq j<s \\
%     0, & \phantom{}1\leq i\leq,~\phantom{:::} j=s
%   \end{cases}
% \end{align*}
Thus, in \eqref{op:bp_tv}, $\left \| D_v x \right \|_1$ (resp.,
$\left \| D_h x \right \|_1$) is the total variation (TV) of the
signal in the vertical (resp., horizontal) direction. The parameters
$\alpha$ and $\beta$ are weighting parameters. The description of BPVV
in \eqref{op:bp_tvW} can be interpreted as changing our assumption of
sparsity in the DCT basis to assuming sparsity in an overcomplete
dictionary \cite{candes_redundant_2011}. \todo{I believe this is
  correct: Sean/Yufan can you confirm or reject?} When written as
\eqref{op:bp_tvW}, the problem is in a form that may be solved with
NESTA~\cite{becker_nesta_2011}.

@@

A simulation comparison between BPDN and BPVV is shown in
Fig.~\ref{fig:BPTV_demonstration}. The first column shows two raster
scans: the top is 512 by 512 pixel image of a CS20NG grating acquired
with an Agilent 5400 AFM and the bottom is a 256 by 256 pixel image of
DNA acquired with an Agilent 5500 AFM. For the CS20NG grating, we
sampled 15\% of the pixels with 50 pixel $\mu$-paths. We sampled 25\%
of the DNA image with 25 pixel long $\mu$-paths. The second columns
shows the full frame reconstructions using BPDN and the third columns
shows the result using BPVV. The fourth column zooms in on the region
indicated by the red box in column one. Similarly, the fifth and six
columns show the detail of the reconstructions. The reconstruction
results show the artifacts apparent in the BPDN reconstruction are
largely mitigated using BPVV. Thus, in the remainder of the paper, we
focus exclusively on BPVV. As for reconstruction time, it took about
1.5 seconds to reconstruct the DNA image and about 7 seconds for the
CS20NG.

@@

Through a breakout box, the Agilent 5400 provides access to the
$Z$-axis deflection signal. When the Agilent software is set to
open-loop mode, a $\pm10$~volt input on the standard control box
allows control of the $Z$-axis of the N9524a piezo scanner, which has
a total range of 7 $\mu$m. Initial (course) engagement of the tip to
the sample is performed using the PicoView software before control is
handed over to the custom controllers.

All control logic is programmed using LabVIEW 2019 and compiled to a
Xilinx Spartan-6 LX150 Field Programmable Gate Array (FPGA) inside the
cRIO-9082. The cRIO includes a 16-Bit, 100 kHz NI-9215
analog-to-digital input module and a 16-Bit, 100 kHz NI-9263
digital-to-analog output module. All control loops are implemented
using a 25 kHz sampling rate.

@@

Transition between, and operation in, these different stages is
implemented as a simple state machine, summarized in
Table~\ref{tab:cs_tasks}. The main job of the state machine is to
adjust the reference signals to accomplish each task. In the
following, $r_X$ and $r_Y$ and $r_Z$ refer to the references for the
$X$ and $Y$ axes while $r_{Z,s}$ and $\rzup$ refer to the $Z$-axis
setpoints during scanning and retraction, respectively.

@@

\subsection{Cantilever does not need to fully disengage.}
\todo{Not clear to me yet that friction is useful if I scan parallel
  to cantilever. Also not clear to me that the normal force and shear
  force are orthogonal concepts. Maybe the macroscopic analogy breaks
  down, but thinking of a knife pushed into a surface and then
  dragged...} In our initial implementation
\cite{braker_hardware_2018}, we insisted that the during
tip-retraction (state-5), the cantilever tip should fully break
contact with the surface. Here, we impose no such requirement. Rather,
we change the $Z$-setpoint during tip-retraction to only pull away far
enough that the deflection signal stays low (i.e., below the scanning
setpoint), even while we run across the surface. We choose \rzup so
that the interaction between the tip and sample becomes purely
adhesive. In general, stable imaging in is not possible at \rzup and
complete dis-engagement occasionally occurs. This can be seen in bottom pane of
Fig.~\ref{fig:cs_cycle} during the second $XY$-move (black portion) where the tip detaches partway through the move.
%Fig.~\ref{fig:afm_bd_dinv} which selects either the scanning setpoint $r_{Z,s}$ or the retraction setpoint $r_{Z,\textrm{up}}$ based the current state.
When dis-engagement does occur, the $Z$-axis control loop implements an anti-windup scheme to prevent the control from saturating.

>>>>

\author{Roger A. Braker, \textit{Student Member, IEEE}, Yufan Luo, \textit{Student Member, IEEE} , Lucy Y. Pao, \textit{Fellow, IEEE} and Sean B. Andersson, \textit{Senior Member, IEEE, and Member, ASME}
  \thanks{Braker and Pao are with the Dept. of Electrical, Computer, and
    Energy Engineering at the University of Colorado, 425 UCB,
    Boulder, CO 80309, USA. %Phone: +1 (303) 492-2360. Fax:
    +1 (303) 492-2758.}%R. A. Braker (corresponding author
    %roger.braker@colorado.edu) is a graduate student and L.Y. Pao
    %(pao@colorado.edu) is the Palmer  Professor.}
    \thanks{Luo is with the Division of Systems Engineering and Andersson is with the Division of Systems Engineering and the Department of Mechanical Engineering, Boston University, Boston, MA 02215, USA.}
  \thanks{This work was supported in part by the US National Science
    Foundation (NSF Grants CMMI-1234980, CMMI-1234845, CMMI-1562031, and DBI-1352729)  Agilent Technologies,
    Inc., and the Hanse-Wissenschaftskolleg Institute for Advanced Study in Delmenhorst, Germany.} }

\maketitle
\begin{abstract}
  Undersampling-based approaches in Atomic Force Microscopy (AFM) aim
  to reduce the time to acquire an image by reducing the number of
  measurements needed while still maintaining image quality. The approach consists of two components: data acquisition and image reconstruction. While the underlying idea is straightforward, practical implementation of a working scheme involves solving a variety of non-trivial problems. In this
  paper, we describe one such hardware implementation and demonstration based on the use of collections of short scans
  known as $\mu$--paths. Using a commercial AFM, we acquire data on a
  grating sample. Reconstructions are made from these data using
  % three approaches: (1) inpainting, an interpolation technique that
  % diffuses information from sampled locations into unsampled
  % regions, (2)
  % basis-pursuit, a compressive sensing-based algorithm that seeks to
  % minimize a
  % measure of sparsity in the underlying image, and (3)
  a new variant of basis pursuit, Basis Pursuit with Vertical
  Variation, that is designed to reduce artifacts arising from
  the sampling pattern. The quality of the resulting images is
  compared to images from standard raster scans of the same regions at
  comparable imaging rates using both the peak signal-to-noise ratio
  and the structured similarity index metric and a new metric we call
  the relative damage index. These images are compared both to standard raster scans as well as to simple subline sampling where only a subset of the rows of the image are acquired (followed by reconstruction). \textcolor{red}{These experiments demonstrate that at slow scan
  rates, the $\mu-$path scheme produces similar image quality with
  significantly less sampling (and thus less tip-sample interaction)
  and that at high rates, the undersampling scheme produces higher
  quality images than raster scanning.}
\end{abstract}

@@

In addition to imaging static samples, AFM is used to study dynamics
in systems with nanometer-scale features
\cite{Yang:2017im,Shibata:2017da,Shibata:2015jd,Ando:2014ja}. The
image acquisition time of conventional AFMs, however, is typically on
the order of seconds to minutes, severely limiting the time-scales
that can be explored. Driven by the need for faster imaging, there
continue to be many active research efforts to overcome this
challenge. Since an AFM is fundamentally a mechanical microscope, many
of the approaches to high-speed AFM (HS-AFM) have focused on
modifications to the physical components. These have included the use
of small, fast cantilevers \cite{Viani:1999tp,
  braunsmann2010high,Adams:2016hg}, the development of faster
actuators \cite{Maroufi:2015gt,Yong:2012kd,Kenton:2012cm}, and
the application of advanced controllers
\cite{Rana:2018es,Yong:2015gr,butterworth2010adaptive,salapaka2002high}.
By combining different techniques, there are current generation,
high-end instruments that can image at rates on the order of 1-10
frames/sec \cite{Ando:2014ja}. However, the fastest rates are achieved
only over small scan ranges and there remain many systems of interest
whose dynamics are faster even than these instruments can achieve. In
addition, there is a large installed base of much slower AFMs that can
benefit from alternative approaches to improving imaging rate.

A complimentary class of HS-AFM approaches considers modifications to
the sampling scheme rather than the system mechanics. Alternative scan
trajectories like spirals, cycloids, and Lissajous figures, have been
used in place of the standard raster pattern. While these trajectories are
easier for the actuators to follow, allowing the tip to be moved more
quickly and leading to faster image acquisition on an otherwise
unmodified instrument
\cite{Bazaei:2017dm,Wu:2015dt,Rana:2014bj,Tuma:2012hv,Yong:2010gm}, they are ultimately limited by the vertical bandwidth of
the instrument for regulating the deflection signal \cite{Teo:2016ev}.

A different use of non-raster scan patterns are those that seek to improve imaging rate by reducing sampling rather than increasing speed. This class includes local scanning methods that use the measurements in real-time to steer the tip to focus the scan on features of
interest \cite{Hartman:2018ja,Wen:2018fl,Zhang:2015cf,Huang:2014dw}.
While these have been shown to yield an order of magnitude or better
improvement in imaging rate, they are limited in the class of samples
that can be imaged and do not produce a full image with all the corresponding context.

@@

One simple way to create undersampling schemes is by modifying
existing full scan patterns including raster, spiral, and Lissajous
scanning. For example, subline sampling is generated by randomly
skipping some of the horizontal lines in raster scanning
\cite{han2018reconstruction,chen2013enhancement}. For spiral and
Lissajous patterns, the scan parameters (frequency and amplitude) can
be selected to ensure the trajectory only passes through a desired
fraction of the pixels in the final image. The scanning time for these
smooth undersampling patterns can be estimated based on the proportion
of the pixels in the trajectory. However, results in the literature of
CS make clear that \textit{randomness} in the sampling pattern is
essential for creating good reconstructions in the general case. %The smooth nature of these patterns reduces their randomness and thus leads to less accurate reconstructions than a completely random sampling of the pixels for the same sampling fraction.

In AFM, implementing a truly random pattern requires that the tip be
engaged with the surface to collect a measurement, lifted and moved to
the new location, and engaged again.
Because the re-engagement process
is typically slow, this can lead to excessively long image acquisition
times and negate the gains from undersampling \cite{andersson2012non}.
In \cite{maxwell2014compressed}, we introduced an undersampling scheme,
called a $\mu$-path pattern, which consists of short randomly placed,
horizontal scans (illustrated in Fig.~\ref{fig:mu_mask}). It
is designed to balance randomness (to ensure good
reconstruction) with continuous scanning (to reduce the number of tip
engagements).

Our previous work on the $\mu$-path pattern, using theoretical
calculations, simulations, and a preliminary implementation,
demonstrates that the approach can achieve scanning time
reduction while maintaining faithful image reconstruction
\cite{maxwell2014compressed,Luo:2015tu, braker_hardware_2018}. 
The main contributions of this paper relative to that past work are as follows.
\begin{itemize}
\item We introduce a new reconstruction algorithm which is designed to
minimize the artifacts arising from the structure of the $\mu$-path
scanning pattern. This algorithm is described and compared to existing
reconstruction methods through simulations in Sec. \ref{sec:reconstructionMethods}.
\item In our prior experimental prototype \cite{braker_hardware_2018}, we used very basic integral controllers for all three axes for both raster and $\mu$-path scans. Here, we design compensators for each axis using standard loop-shaping techniques (Sec.~\ref{sec:control}), which result in a substantial increase in closed-loop bandwidth (e.g., 50 Hz to 450~Hz for the $Z$-axis).
We show in Sec.~\ref{sec:ff_control} that one of the limitations of $\mu$-path scanning is the discontinuous $X$-axis reference (e.g., in moving between $\mu$-paths) which excites the cross-coupling modes between the $X$ and $Z$ axes, an effect which neither our prior experimental work nor simulation studies accounted for. We mitigate the cross coupling with a feed-forward design.
\item We experimentally compare (Sec. \ref{sec:results}) $\mu$-path scanning to not only full density raster scans but also to coarse raster scans which are interpolated to a full image. This is in contrast to typical sub-line sampling which takes scan lines at random. The clear advantage to a coarse raster scan is that it requires no modification to the instrument, only to the post-processing.
\item The bulk of this work is based on experiment, not simulation, in contrast to prior work by ourselves \cite{luo2015comparison,maxwell2014compressed} and others \cite{han_optimal_2018,jensen_reconstruction_2013,oxvig_structure_2017}. Our main focus here is on delineating the practical issues which must be addressed to effectively realize $\mu$-path scanning, which has only seen limited attention in the past.
\end{itemize}

@@

\noindent
\begingroup \setlength{\tabcolsep}{1pt}
\begin{figure*}[ht!]
  \centering
  \begin{tabular}{cccccc}
    \textit{\small original image} & \textit{\small BPDN} & \textit{\small BPVV}
    & \textit{\small detail of original} & \textit{\small detail of BPDN} & \textit{\small detail of BPVV} \\
    \includesvg[width=0.16\textwidth]{figures/cs20ng_og.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_og_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp_zoom.svg}
    %
    &   \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv_zoom.svg}\\
    % 
    % ------------------------------
    % Row
    % 2
    % --------------------------------------------
    \includesvg[width=0.16\textwidth]{figures/dna_og.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bp.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_og_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bp_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bptv_zoom.svg}
  \end{tabular}
  \caption{Reconstruction comparison between BPDN and BPVV. (first
    column) Original raster-scanned image. (second column) BPDN
    reconstruction from random, \textcolor{red}{40 pixel long $\mu$-paths with 25\%
    sampling}. (third column) BPVV reconstruction from the same
    sub-sampled data. (remaining columns) Corresponding details from
    the red boxes indicated in the raster image. The results show that
    BPVV reduces the artifacts arising from the horizontal scans of
    the $\mu$-path pattern that appear in BPDN reconstructions.}
  \label{fig:BPTV_demonstration}
\end{figure*}
\endgroup

In the AFM application, the probe can only measure a single pixel at a
time. Thus the rows of $R$ are a subset of those of an $n$ by $n$
identity matrix. Ideally, the sparsity basis and the measurement
matrix $R$ will have a low \textit{mutual coherence}, a measure of how
each of the rows of $R$ (the measurements) ``spreads out'' in the
domain of $U$ \cite{candes2007sparsity}. In the following, we assume
that $U^TU=I$, and in the simulation and experimental results we make the specific choice of $U^T$ as the Discrete Cosine Transform (DCT). The DCT basis
generally keeps a good balance between achieving a low mutual
coherence between $U$ and the $\mu$-path sensing matrices $R$ and
providing a high sparsity for typical AFM sample images.

Basis pursuit with denoising (BPDN) is one common realization of the
CS-based reconstruction problem, and is given by the optimization
\begin{equation}
  \min_{x} \left \| U^Tx \right \|_1 \quad
  \text{subj. to}\quad %x\in Q_p = \{x:~\|Rx - y\|_2 < \sigma\} 
  \|Rx-y\|_2 < \sigma \label{op:bp}
\end{equation}
where $\sigma$ represents uncertainty in the measurements. BPDN
essentially searches for the sparsest signal among the candidates
that match the measurements to within the tolerance $\sigma$.

@@

To mitigate these affects, we propose a new variant of BPDN that we
term basis pursuit with vertical variation (BPVV). BPVV adds a
vertical total variation penalty in the spatial domain to the
optimization objective. That is, \eqref{op:bp} is modified to \todo{It
  has been my observation that some benefit is gained by including
  $D_h$, but weighting it less than $D_v$. Not sure if we should leave
  it in. If so, BPVV should be BPTV or something.}
\begin{align}
  \min_{x\in Q_p}~f(x) &= \min_{x\in Q_p} \left\|U^Tx\right\|_1
                         + \alpha\left\|D_vx\right\|_1
                         + \beta\left\|D_hx\right\|_1 \label{op:bp_tv}\\
                       &=\min_{x\in Q_p} \left\|W^Tx\right\|_1 \label{op:bp_tvW}
                       % \quad W
                       % = \begin{bmatrix}U&\alpha_hD^T_h&\alpha_vD^T_v \end{bmatrix}^T
                       % \nonumber
\end{align}
where
\begin{equation*}
  W = \begin{bmatrix}U^T&\alpha D_v&\beta D_h \end{bmatrix}
\end{equation*}
with ${D_v x=\text{vec}(\nabla_v X)}$ (resp.,
${D_h x=\text{vec}(\nabla_h X)}$) representing a discrete gradient of
a matrix in the vertical (resp., horizontal) direction (see, e.g.,
\cite[Sec. 6.1,]{becker_nesta_2011}). 
% and
% \begin{align*}
%   (\nabla_v X)_{i,j} = 
%   \begin{cases}
%     X_{i+1,j} - X_{i,j}, & 1\leq i<h,~ 1\leq j\leq s,\\
%     0, & \phantom{:}i=h,~\phantom{:::::} 1\leq j\leq s.
%   \end{cases}\\
%   (\nabla_h X)_{i,j} =
%   \begin{cases}
%     X_{i,j+1} - X_{i,j}, & 1\leq i\leq h,~1\leq j<s \\
%     0, & \phantom{}1\leq i\leq,~\phantom{:::} j=s
%   \end{cases}
% \end{align*}
Thus, in \eqref{op:bp_tv}, $\left \| D_v x \right \|_1$ (resp.,
$\left \| D_h x \right \|_1$) is the total variation (TV) of the
signal in the vertical (resp., horizontal) direction. The parameters
$\alpha$ and $\beta$ are weighting parameters. The description of BPVV
in \eqref{op:bp_tvW} can be interpreted as changing our assumption of
sparsity in the DCT basis to assuming sparsity in an overcomplete
dictionary \cite{candes_redundant_2011}. \todo{I believe this is
  correct: Sean/Yufan can you confirm or reject?} When written as
\eqref{op:bp_tvW}, the problem is in a form that may be solved with
NESTA~\cite{becker_nesta_2011}. Note that the BPVV designation technically holds only for $\beta=0$; our experience, however, shows that including a small but nonzero $\beta$ (relative to $\alpha$) improves reconstruction.

@@

A simulation comparison between BPDN and BPVV is shown in
Fig.~\ref{fig:BPTV_demonstration}. The first column shows two raster
scans: the top is a 512 by 512 pixel image of a CS20NG grating (Ted Pella, USA) acquired
with an Agilent 5400 AFM and the bottom is a 256 by 256 pixel image of
DNA acquired with an Agilent 5500 AFM. For the CS20NG grating, we
sampled 15\% of the pixels with 50 pixel $\mu$-paths and 25\%
of the DNA image with 25 pixel long $\mu$-paths. The second columns
shows the full frame reconstructions using BPDN and the third columns
shows the result using BPVV (\textcolor{red}{with $\alpha = ??$ and $\beta = ??$}). The fourth column zooms in on the region
indicated by the red box in column one. Similarly, the fifth and six
columns show the detail of the reconstructions. The reconstruction
results show the artifacts apparent in the BPDN reconstruction are
largely mitigated using BPVV. Thus, in the remainder of the paper, we
focus exclusively on BPVV. These reconstructions wth BPVV took about
1.5 seconds for the DNA image and about 7 seconds for the
CS20NG (on a typical laptop computer).

@@

Through a breakout box, the Agilent 5400 provides access to the
$Z$-axis deflection signal. When the Agilent software is set to
open-loop mode, a $\pm10$~volt input on the standard control box
allows control of the $Z$-axis of the N9524a piezo scanner, which has
a total range of 7 $\mu$m. Initial (coarse) engagement of the tip to
the sample is performed using the PicoView software before control is
handed over to the custom controllers.

All control logic is programmed using LabVIEW 2019 and compiled to a
Xilinx Spartan-6 LX150 Field Programmable Gate Array (FPGA) inside the
cRIO-9082. The cRIO includes 16-bit, 100 kHz analog-to-digital input (NI-9215)) and digital-to-analog output (NI-9263) modules.
%a 16-Bit, 100 kHz NI-9215analog-to-digital input module and a 16-Bit, 100 kHz NI-9263 digital-to-analog output module. 
All control loops are implemented
using a 25 kHz sampling rate.

@@

Transition between, and operation in, these different stages is
implemented as a simple state machine, summarized in
Table~\ref{tab:cs_tasks}. The main job of the state machine is to
adjust the reference signals to accomplish each task. In the
following, $r_X$ and $r_Y$ refer to the references for the
$X$ and $Y$ axes while $r_{Z,s}$ and $\rzup$ refer to the $Z$-axis
setpoints during scanning and retraction, respectively.

@@

\subsection{Cantilever does not need to fully disengage.}
\todo{Not clear to me yet that friction is useful if I scan parallel
  to cantilever. Also not clear to me that the normal force and shear
  force are orthogonal concepts. Maybe the macroscopic analogy breaks
  down, but thinking of a knife pushed into a surface and then
  dragged...}

In order to minimize the scanning time, tip tip should be moved as fast as possible when trainsitioning between $\mu$-path locations (state (1)). Since such speeds can be beyond the bandwidth of the $z-$axis controller, it is important to reduce the tip-sample interaction force to prevent tip sear and sample damage. In our initial implementation, this was achieved by 
%In our initial implementation
\cite{braker_hardware_2018}, we insisted that the during
tip-retraction (state (5)), the cantilever tip should fully break
contact with the surface. Here, we impose no such requirement. Rather,
we change the $Z$-setpoint during tip-retraction to only pull away far
enough that the deflection signal stays low (i.e., below the scanning
setpoint), even while we run across the surface. We choose \rzup so
that the interaction between the tip and sample becomes purely
adhesive. In general, stable imaging is not possible at \rzup and
complete dis-engagement occasionally occurs. This can be seen in bottom pane of
Fig.~\ref{fig:cs_cycle} during the second $XY$-move (black portion) where the tip detaches partway through the move.
%Fig.~\ref{fig:afm_bd_dinv} which selects either the scanning setpoint $r_{Z,s}$ or the retraction setpoint $r_{Z,\textrm{up}}$ based the current state.
When dis-engagement does occur, the $Z$-axis control loop implements an anti-windup scheme to prevent the control from saturating.
